---
title: "Manuscript_NestSuccess"
author: "Madeline Aberg"
date: '2022-07-02'
output: pdf_document
---
# HOLA & LBCU nest success analysis
# 1 May 2024


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(lubridate)
library(lme4)
library(MuMIn)
library(car)
library(jtools)
library(MASS)
library(ggplot2)
```

# Load Data
HOLA
```{r}
HOLAnests.Active <- read.csv( "C:/Users/Madeline/Desktop/EcologicalImpactsManuscript/CleanData/HOLA_NestSuccess_Clean_29April2024.csv")

HOLAnests.Veg <- read.csv("C:/Users/Madeline/Desktop/EcologicalImpactsManuscript/CleanData/HOLA_NestSuccess_Veg_Clean_29April2024.csv")
```

LBCU
```{r}
LBCUnests <- read.csv("C:/Users/Madeline/Desktop/EcologicalImpactsManuscript/CleanData/LBCU_NestSuccess_Clean_29April2024.csv")
```

# Mayfield Nest Success Estimates
## HOLA
### All active nests
```{r}
# Create high and low recreation datasets 
  LowRec <- subset(HOLAnests.Active, Location == "CinderCone")
  HighRec <- subset(HOLAnests.Active, Location == "PleasantValley")
  LowRec2020 <- filter(LowRec, Year == 2020)
  LowRec2021 <- filter(LowRec, Year == 2021)
  HighRec2020 <- filter(HighRec, Year == 2020)
  HighRec2021 <- filter(HighRec, Year == 2021)

# HighRec2020 DSR & Mayfield Est. ---------------------------------------------------------------

    # calculate exposure days for each nest = end_julian - found_julian 
     HighRec2020$exposure_days_total <- HighRec2020$endDOY - HighRec2020$foundDOY

  
    # create a variable with the number of failures
        HighRec2020$fail_yn <- ifelse(HighRec2020$Fate.2 == "failed", 1, 0)
    
        total_fail <- sum(HighRec2020$fail_yn)
        
        total_fail # total nest failures
    
    # create a variable with the total number of exposure days
        exposure_days <- sum(HighRec2020$exposure_days_total)
      
        exposure_days # total exposure days for all nests in the study
    
    # calculate daily failure rate r = Failed/Exposure days
  
        r_high2020H <- total_fail / exposure_days
        
        r_high2020H # daily failure rate for the nests 
    
    # calculate daily survival rate S = 1 - r
      
        S_high2020H <-  1 - r_high2020H 

    # calculate mayfield success = S^d where d = number of days in the nesting cycle 

        d_H <- 19 
        MS_high2020H <- S_high2020H ^ d_H

        MS_high2020H

 # LowRec2020 DSR & Mayfield Est. ---------------------------------------------------------------

    # calculate exposure days for each nest = end_julian - found_julian 
     LowRec2020$exposure_days_total <- LowRec2020$endDOY - LowRec2020$foundDOY
    
    # create a variable with the number of failures
        LowRec2020$fail_yn <- ifelse(LowRec2020$Fate.2 == "failed", 1, 0)
    
        total_fail <- sum(LowRec2020$fail_yn)
        
        total_fail # total nest failures
    
    # create a variable with the total number of exposure days
        exposure_days <- sum(LowRec2020$exposure_days_total)
      
        exposure_days # total exposure days for all nests in the study
    
    # calculate daily failure rate r = Failed/Exposure days
  
        r_low2020H <- total_fail / exposure_days
        
        r_low2020H # daily failure rate for the nests 
    
    # calculate daily survival rate S = 1 - r
      
        S_low2020H <-  1 - r_low2020H 

    # calculate mayfield success = S^d where d = number of days in the nesting cycle 
        MS_low2020H <- S_low2020H ^ d_H

        MS_low2020H      
        
 # HighRec2021 DSR & Mayfield Est. ---------------------------------------------------------------

    # calculate exposure days for each nest = end_julian - found_julian 
     HighRec2021$exposure_days_total <- HighRec2021$endDOY - HighRec2021$foundDOY

  
    # create a variable with the number of failures
        HighRec2021$fail_yn <- ifelse(HighRec2021$Fate.2 == "failed", 1, 0)
    
        total_fail <- sum(HighRec2021$fail_yn)
        
        total_fail # total nest failures
    
    # create a variable with the total number of exposure days
        exposure_days <- sum(HighRec2021$exposure_days_total)
      
        exposure_days # total exposure days for all nests in the study
    
    # calculate daily failure rate r = Failed/Exposure days
  
        r_high2021H <- total_fail / exposure_days
        
        r_high2021H # daily failure rate for the nests 
    
    # calculate daily survival rate S = 1 - r
      
        S_high2021H <-  1 - r_high2021H 

    # calculate mayfield success = S^d where d = number of days in the nesting cycle 

        d_H <- 19 
        MS_high2021H <- S_high2021H ^ d_H

        MS_high2021H   
        
        
 # LowRec2021 DSR & Mayfield Est. ---------------------------------------------------------------

    # calculate exposure days for each nest = end_julian - found_julian 
     LowRec2021$exposure_days_total <- LowRec2021$endDOY - LowRec2021$foundDOY
    
    # create a variable with the number of failures
        LowRec2021$fail_yn <- ifelse(LowRec2021$Fate.2 == "failed", 1, 0)
    
        total_fail <- sum(LowRec2021$fail_yn)
        
        total_fail # total nest failures
    
    # create a variable with the total number of exposure days
        exposure_days <- sum(LowRec2021$exposure_days_total)
      
        exposure_days # total exposure days for all nests in the study
    
    # calculate daily failure rate r = Failed/Exposure days
  
        r_low2021H <- total_fail / exposure_days
        
        r_low2021H # daily failure rate for the nests 
    
    # calculate daily survival rate S = 1 - r
      
        S_low2021H <-  1 - r_low2021H 

    # calculate mayfield success = S^d where d = number of days in the nesting cycle 
        MS_low2021H <- S_low2021H ^ d_H

        MS_low2021H     
```

### LBCU?
```{r}
# Create high and low recreation datasets -----------------------------------------------
  LowRec <- subset(HOLAnests.Veg, Location == "CinderCone")
  HighRec <- subset(HOLAnests.Veg, Location == "PleasantValley")
  LowRec2020 <- filter(LowRec, Year == 2020)
  LowRec2021 <- filter(LowRec, Year == 2021)
  HighRec2020 <- filter(HighRec, Year == 2020)
  HighRec2021 <- filter(HighRec, Year == 2021)

# HighRec2020 DSR & Mayfield Est. ---------------------------------------------------------------

    # calculate exposure days for each nest = end_julian - found_julian 
     HighRec2020$exposure_days_total <- HighRec2020$endDOY - HighRec2020$foundDOY

  
    # create a variable with the number of failures
        HighRec2020$fail_yn <- ifelse(HighRec2020$Fate.2 == "failed", 1, 0)
    
        total_fail <- sum(HighRec2020$fail_yn)
        
        total_fail # total nest failures
    
    # create a variable with the total number of exposure days
        exposure_days <- sum(HighRec2020$exposure_days_total)
      
        exposure_days # total exposure days for all nests in the study
    
    # calculate daily failure rate r = Failed/Exposure days
  
        r_high2020H <- total_fail / exposure_days
        
        r_high2020H # daily failure rate for the nests 
    
    # calculate daily survival rate S = 1 - r
      
        S_high2020H <-  1 - r_high2020H 

    # calculate mayfield success = S^d where d = number of days in the nesting cycle 

        d_H <- 19 
        MS_high2020H <- S_high2020H ^ d_H

        MS_high2020H

 # LowRec2020 DSR & Mayfield Est. ---------------------------------------------------------------

    # calculate exposure days for each nest = end_julian - found_julian 
     LowRec2020$exposure_days_total <- LowRec2020$endDOY - LowRec2020$foundDOY
    
    # create a variable with the number of failures
        LowRec2020$fail_yn <- ifelse(LowRec2020$Fate.2 == "failed", 1, 0)
    
        total_fail <- sum(LowRec2020$fail_yn)
        
        total_fail # total nest failures
    
    # create a variable with the total number of exposure days
        exposure_days <- sum(LowRec2020$exposure_days_total)
      
        exposure_days # total exposure days for all nests in the study
    
    # calculate daily failure rate r = Failed/Exposure days
  
        r_low2020H <- total_fail / exposure_days
        
        r_low2020H # daily failure rate for the nests 
    
    # calculate daily survival rate S = 1 - r
      
        S_low2020H <-  1 - r_low2020H 

    # calculate mayfield success = S^d where d = number of days in the nesting cycle 
        MS_low2020H <- S_low2020H ^ d_H

        MS_low2020H      
        
        
# HighRec2021 DSR & Mayfield Est. ---------------------------------------------------------------

    # calculate exposure days for each nest = end_julian - found_julian 
     HighRec2021$exposure_days_total <- HighRec2021$endDOY - HighRec2021$foundDOY

  
    # create a variable with the number of failures
        HighRec2021$fail_yn <- ifelse(HighRec2021$Fate.2 == "failed", 1, 0)
    
        total_fail <- sum(HighRec2021$fail_yn)
        
        total_fail # total nest failures
    
    # create a variable with the total number of exposure days
        exposure_days <- sum(HighRec2021$exposure_days_total)
      
        exposure_days # total exposure days for all nests in the study
    
    # calculate daily failure rate r = Failed/Exposure days
  
        r_high2021H <- total_fail / exposure_days
        
        r_high2021H # daily failure rate for the nests 
    
    # calculate daily survival rate S = 1 - r
      
        S_high2021H <-  1 - r_high2021H 

    # calculate mayfield success = S^d where d = number of days in the nesting cycle 
        MS_high2021H <- S_high2021H ^ d_H

        MS_high2021H
        
 # LowRec2021 DSR & Mayfield Est. ---------------------------------------------------------------

    # calculate exposure days for each nest = end_julian - found_julian 
     LowRec2021$exposure_days_total <- LowRec2021$endDOY - LowRec2021$foundDOY
    
    # create a variable with the number of failures
        LowRec2021$fail_yn <- ifelse(LowRec2021$Fate.2 == "failed", 1, 0)
    
        total_fail <- sum(LowRec2021$fail_yn)
        
        total_fail # total nest failures
    
    # create a variable with the total number of exposure days
        exposure_days <- sum(LowRec2021$exposure_days_total)
      
        exposure_days # total exposure days for all nests in the study
    
    # calculate daily failure rate r = Failed/Exposure days
  
        r_low2021H <- total_fail / exposure_days
        
        r_low2021H # daily failure rate for the nests 
    
    # calculate daily survival rate S = 1 - r
      
        S_low2021H <-  1 - r_low2021H 

    # calculate mayfield success = S^d where d = number of days in the nesting cycle 
        MS_low2021H <- S_low2021H ^ d_H

        MS_low2021H     
```

## LBCU
```{r}
# Create high and low recreation datasets -----------------------------------------------
  LowRec <- subset(LBCUnests, Subsite == "Cinder Cone")
  HighRec <- subset(LBCUnests, Subsite == "Pleasant Valley")
  LowRec2019 <- filter(LowRec, year == 2019)
  LowRec2020 <- filter(LowRec, year == 2020)
  LowRec2021 <- filter(LowRec, year == 2021)
  HighRec2019 <- filter(HighRec, year == 2019)
  HighRec2020 <- filter(HighRec, year == 2020)
  HighRec2021 <- filter(HighRec, year == 2021)


d_L = 33 # length of nesting cycle

# HighRec2019 DSR & Mayfield Est. ---------------------------------------------------------------

    # calculate exposure days for each nest = end_julian - found_julian 
     exposure_days_total <- sum(HighRec2019$exposure_days)

    # create a variable with the number of failures
        HighRec2019 <- mutate(HighRec2019, fail_yn = case_when(
          fate == "fail" | fate == "Fail" ~ 1, 
          fate == "hatch" | fate == "Hatch" ~0))
      
        total_fail <- sum(HighRec2019$fail_yn)
        
        total_fail # total nest failures
  
    
    # calculate daily failure rate r = Failed/Exposure days
  
        r_high2019L <- total_fail / exposure_days_total
        
        r_high2019L  # daily failure rate for the nests 
    
    # calculate daily survival rate S = 1 - r
      
        S_high2019L <-  1 - r_high2019L 

    # calculate mayfield success = S^d where d = number of days in the nesting cycle 
        MS_high2019L <- S_high2019L ^ d_L

        MS_high2019L
        
# HighRec2020 DSR & Mayfield Est. ---------------------------------------------------------------

    # calculate exposure days for each nest = end_julian - found_julian 
     exposure_days_total <- sum(HighRec2020$exposure_days)

    # create a variable with the number of failures
        HighRec2020 <- mutate(HighRec2020, fail_yn = case_when(
          fate == "fail" | fate == "Fail" ~ 1, 
          fate == "hatch" | fate == "Hatch" ~0))
      
        total_fail <- sum(HighRec2020$fail_yn)
        
        total_fail # total nest failures
    
    
    # calculate daily failure rate r = Failed/Exposure days
  
        r_high2020L <- total_fail / exposure_days_total
        
        r_high2020L # daily failure rate for the nests 
    
    # calculate daily survival rate S = 1 - r
      
        S_high2020L <-  1 - r_high2020L 

    # calculate mayfield success = S^d where d = number of days in the nesting cycle 
        MS_high2020L <- S_high2020L ^ d_L

        MS_high2020L
        
# HighRec2021 DSR & Mayfield Est. ---------------------------------------------------------------

    # calculate exposure days for each nest = end_julian - found_julian 
     exposure_days_total <- sum(HighRec2021$exposure_days)

    # create a variable with the number of failures
        HighRec2021 <- mutate(HighRec2021, fail_yn = case_when(
          fate == "fail" | fate == "Fail" ~ 1, 
          fate == "hatch" | fate == "Hatch" ~0))
      
        total_fail <- sum(HighRec2021$fail_yn)
        
        total_fail # total nest failures
    
    # calculate daily failure rate r = Failed/Exposure days
  
        r_high2021L <- total_fail / exposure_days_total
        
        r_high2021L # daily failure rate for the nests 
    
    # calculate daily survival rate S = 1 - r
      
        S_high2021L <-  1 - r_high2021L 

    # calculate mayfield success = S^d where d = number of days in the nesting cycle 
        MS_high2021L <- S_high2021L ^ d_L

        MS_high2021L
        
# LowRec2019 DSR & Mayfield Est. ---------------------------------------------------------------

  # calculate exposure days for each nest = end_julian - found_julian 
     exposure_days_total <- sum(LowRec2019$exposure_days)

  # create a variable with the number of failures
        LowRec2019 <- mutate(LowRec2019, fail_yn = case_when(
          fate == "fail" | fate == "Fail" ~ 1, 
          fate == "hatch" | fate == "Hatch" ~0))
      
        total_fail <- sum(LowRec2019$fail_yn)
        
        total_fail # total nest failures
    
    # calculate daily failure rate r = Failed/Exposure days
  
        r_low2019L <- total_fail / exposure_days_total
        
        r_low2019L # daily failure rate for the nests 
    
    # calculate daily survival rate S = 1 - r
      
        S_low2019L <-  1 - r_low2019L 

    # calculate mayfield success = S^d where d = number of days in the nesting cycle 
        MS_low2019L <- S_low2019L ^ d_L
        MS_low2019L
        
# LowRec2020 DSR & Mayfield Est. ---------------------------------------------------------------

  # calculate exposure days for each nest = end_julian - found_julian 
      exposure_days_total <- sum(LowRec2020$exposure_days)

  # create a variable with the number of failures
        LowRec2020 <- mutate(LowRec2020, fail_yn = case_when(
          fate == "fail" | fate == "Fail" ~ 1, 
          fate == "hatch" | fate == "Hatch" ~0))
      
        total_fail <- sum(LowRec2020$fail_yn)
        
        total_fail # total nest failures
    
    # calculate daily failure rate r = Failed/Exposure days
  
        r_low2020L <- total_fail / exposure_days_total
        
        r_low2020L # daily failure rate for the nests 
    
    # calculate daily survival rate S = 1 - r
      
        S_low2020L <-  1 - r_low2020L 

    # calculate mayfield success = S^d where d = number of days in the nesting cycle 
        MS_low2020L <- S_low2020L ^ d_L
        MS_low2020L
        
# LowRec2021 DSR & Mayfield Est. ---------------------------------------------------------------

  # calculate exposure days for each nest = end_julian - found_julian 
      exposure_days_total <- sum(LowRec2021$exposure_days)

  # create a variable with the number of failures
        LowRec2021 <- mutate(LowRec2021, fail_yn = case_when(
          fate == "fail" | fate == "Fail" ~ 1, 
          fate == "hatch" | fate == "Hatch" ~0))
      
        total_fail <- sum(LowRec2021$fail_yn)
        
        total_fail # total nest failures

    # calculate daily failure rate r = Failed/Exposure days
  
        r_low2021L <- total_fail / exposure_days_total
        
        r_low2021L # daily failure rate for the nests 
    
    # calculate daily survival rate S = 1 - r
      
        S_low2021L <-  1 - r_low2021L 

    # calculate mayfield success = S^d where d = number of days in the nesting cycle 
        MS_low2021L <- S_low2021L ^ d_L
        MS_low2021L
```

## Mayfield sample size
```{r}
table(LBCUnests$year, LBCUnests$Subsite)
```


# Nest Success Modeling
## Logistic Exposure Link
```{r}
# Logistic Exposure Link
logexp <- function(days = 1)
{
  linkfun <- function(mu) qlogis(mu^(1/days))
  linkinv <- function(eta) plogis(eta)^days
  mu.eta  <- function(eta) days * plogis(eta)^(days-1) *
    binomial()$mu.eta(eta)
  valideta <- function(eta) TRUE
  link <- paste0("logexp(", days, ")")
  structure(list(linkfun = linkfun, linkinv = linkinv,
                 mu.eta = mu.eta, valideta = valideta, name = link),
            class = "link-glm")
}
```

## HOLA
### predictors
```{r}
# Site-level predictors
prednames_hola <- c("rec", "InitiationDOY", "conceal_sum", "predator", "consp_total")

# loop over each to create histograms for each predictor:
for( p in 1:length(prednames_hola) ){
  # create an object with the ggplot so that you can display it 
  # in a loop 
  a <- ggplot( HOLAnests.Active ) + #choose your data
    theme_bw( base_size = 15 ) + #choose a preset theme
    labs( x = prednames_hola[p] ) + #label x axis using our predictor names
    geom_histogram( aes( get(prednames_hola[p]) ), bins= 10 ) #plot histogram
  # display your plot object
  print( a )
}

# correlation among predictors. > 0.8 would be severe multicolinearity, 0.6 is iffy 
hola.cor <- cor( HOLAnests.Active[ , prednames_hola] )
hola.cor
symnum(hola.cor)

HOLAnests.Veg$orientYN <- as.factor(HOLAnests.Veg$orientYN)
HOLAnests.Veg$Year<- as.factor(HOLAnests.Veg$Year)
```

### variable summary
#### HOLA
```{r}
min.rec <- min(HOLAnests.Veg$rec)
max.rec <- max(HOLAnests.Veg$rec)
mean.rec <- mean(HOLAnests.Veg$rec)
sd.rec <- sd(HOLAnests.Veg$rec)

min.date <- min(HOLAnests.Veg$InitiationDOY)
max.date <- max(HOLAnests.Veg$InitiationDOY)
mean.date <- mean(HOLAnests.Veg$InitiationDOY)
sd.date <- sd(HOLAnests.Veg$InitiationDOY)
```

### model
```{r}
hola.mod <- glm(surv ~ rec_std + initiate_std + orientYN + Year, family = binomial(link = logexp(days = HOLAnests.Veg$ExposureDays)), data = HOLAnests.Veg)

summary(hola.mod)

vif(hola.mod)

library(caret)
varImp(hola.mod)

confint(hola.mod)
```

### in text
```{r}
coef(hola.mod)
plogis(coef(hola.mod))
plogis(0.4249) # intercept se
# effect of 2021
1 - 0.4258462

# effect of orientation
1 - 0.5483371

# effect of rec
a <- 1.79 # intercept
b1 <- -0.42 # rec
# all others at mean (= 0)
rec.min.st <- (min.rec-mean.rec)/(2*sd.rec)
rec.max.st <- (max.rec-mean.rec)/(2*sd.rec)
plogis(a + b1*rec.min.st) - plogis(a + b1*rec.max.st)

# effect of initiation date
a <- 1.79 # intercept
b1 <- -0.4731 # initiation
# all others at mean (= 0)
date.min.st <- (min.date-mean.date)/(2*sd.date)
date.max.st <- (max.date-mean.date)/(2*sd.date)
plogis(a + b1*date.min.st) - plogis(a + b1*date.max.st)
```


## LBCU
```{r}
LBCUnests$rec_std <- stdize(LBCUnests$NESTrec)
LBCUnests$initiate_std <- stdize(LBCUnests$initiationDOY)
LBCUnests$conceal_std <- stdize(LBCUnests$conceal_sum)
LBCUnests$consp_std <- stdize(LBCUnests$consp_total)
LBCUnests$road_std <- stdize(LBCUnests$RoadDist)

LBCUnests <- LBCUnests %>% mutate(
  Year2021 = case_when(
    year == 2021 ~ 1, 
    year == 2020 | year == 2019 ~ 0))

str(LBCUnests$Year2021)
LBCUnests$Year2021 <- as.factor(LBCUnests$Year2021)
```

### predictors
```{r}
# Site-level predictors
prednames_lbcu <- c("NESTrec", "initiationDOY", "conceal_sum", "RoadDist", "consp_total")

# loop over each to create histograms for each predictor:
for( p in 1:length(prednames_lbcu) ){
  # create an object with the ggplot so that you can display it 
  # in a loop 
  a <- ggplot( LBCUnests ) + #choose your data
    theme_bw( base_size = 15 ) + #choose a preset theme
    labs( x = prednames_lbcu[p] ) + #label x axis using our predictor names
    geom_histogram( aes( get(prednames_lbcu[p]) ), bins= 10 ) #plot histogram
  # display your plot object
  print( a )
}

# correlation among predictors. > 0.8 would be severe, 0.6 is iffy 
lbcu.cor <- cor( LBCUnests[ , prednames_lbcu])
lbcu.cor
symnum(lbcu.cor)
```

### variable summary
```{r}
min.rec <- min(LBCUnests$NESTrec)
max.rec <-max(LBCUnests$NESTrec)
mean.rec <- mean(LBCUnests$NESTrec)
sd.rec <- sd(LBCUnests$NESTrec)

min.date <- min(LBCUnests$initiationDOY)
max.date <- max(LBCUnests$initiationDOY)
mean.date <-mean(LBCUnests$initiationDOY)
```


### model
```{r}
lbcu.mod <-glm(surv ~ rec_std + Year2021 + consp_std, family = binomial(link = logexp(days = LBCUnests$exposure_days)), data = LBCUnests)

summary(lbcu.mod)

#confint(lbcu.mod)

# intercept
plogis(coef(lbcu.mod))
plogis(0.2734)

# effect of rec
a <- 1.6312      # intercept
b1 <- -3.0250 # rec
# all others at mean (= 0)
rec.min.st <- (min.rec-mean.rec)/(2*sd.rec)
rec.max.st <- (max.rec-mean.rec)/(2*sd.rec)
plogis(a + b1*rec.min.st) - plogis(a + b1*rec.max.st)
```


# Figure 
## predictors
```{r}
# Range of values for predictions
rec <- seq(10e-12, 2.5e-6,,200)
Idate  <- seq(91, 192,,200)

# standardize recreation values based on the mean & sd from the data
rec_std <- ((rec) - mean.rec)/ (2*sd.rec)
initiate_std <- ((Idate) - mean.date)/(2*sd.date)
```

## hola plot
```{r}
# rec
NewL.rec <- data.frame(rec_std = rec_std, initiate_std = 0, orientYN = 0, Year = 2021)
NewL.rec$orientYN <- as.factor(NewL.rec$orientYN)
NewL.rec$Year <- as.factor(NewL.rec$Year)
x <- predict.glm(hola.mod, NewL.rec, se.fit = TRUE)

y <- as.numeric(x$fit)
y.se <- as.numeric(x$se.fit)

z <- plogis(y)
z.se <- plogis(y.se)

hola.pred <- as.data.frame(z)
hola.pred$rec <- rec
hola.pred$se <- y.se

hola.pred$x <- 1

A1 <- ggplot(data = hola.pred, mapping = aes(x = rec, y = z, 
                                             color = as.factor(x), 
                                             fill = as.factor(x)))+
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = z - se, ymax = pmin(1,z + se)), alpha = 0.1)+
  ylim(0,1) +
  scale_fill_manual(values = "grey40") +
  scale_color_manual(values="grey40") +
  theme_classic()+ 
  labs(x = "Recreation intensity", y = "Predicted probability of nest survival") +
  theme(legend.position = "none")+
  theme(text=element_text(family="Times New Roman", size=12))


  A1
  
ggsave("C:/Users/Madeline/Desktop/NEWholanest.jpg", width = 4, height = 2.5, units = "in", 
       dpi = 400)

```




## lbcu plot
```{r}
newrec <- data.frame(rec_std = rec_std, consp_std = 0, Year2021 = 0)
newrec$Year2021 <-as.factor(newrec$Year2021)
x <- predict.glm(lbcu.mod, newrec, se.fit = TRUE)

y <- as.numeric(x$fit)
y.se <- as.numeric(x$se.fit)

z <- plogis(y)
z.se <- plogis(y.se)

lbcu.pred <- as.data.frame(z)
lbcu.pred$rec <- rec
lbcu.pred$se <- y.se

lbcu.pred$x <- 1

B1 <- ggplot(data = lbcu.pred, mapping = aes(x = rec, y = z, 
                                             color = as.factor(x), 
                                             fill = as.factor(x)))+
  geom_line(linewidth = 1) +
  geom_ribbon(aes(ymin = z - se, ymax = pmin(1,z + se)), alpha = 0.1)+
  ylim(0,1) +
  scale_fill_manual(values = "grey40") +
  scale_color_manual(values="grey40") +
  theme_classic()+ 
  labs(x = "Recreation intensity", y = "Predicted probability of nest survival") +
  theme(legend.position = "none")+
  theme(text=element_text(family="Times New Roman", size=12))


  B1
  
ggsave("C:/Users/Madeline/Desktop/NEWlbcunest.jpg", width = 4, height = 2.5, units = "in", 
       dpi = 400)


```





y <- as.numeric(x$fit)
y.se <- as.numeric(x$se.fit)
lbcu.pred <- as.data.frame(y)
lbcu.pred$rec <- rec
lbcu.pred$se <-y.se
lbcu.pred$y.scale <- (lbcu.pred$y - min(lbcu.pred$y)) / (max(lbcu.pred$y) - min(lbcu.pred$y)) 

B1 <- ggplot(data = lbcu.pred, mapping = aes(x = rec, y = y))+
  geom_line() +
  geom_ribbon(aes(ymin = y - se, ymax = y + se), alpha = 0.1)+
  theme_classic()+ 
  labs(title = "Long-billed curlews - Nesting success", x = "Recreation intensity", y = "Predicted probability of nest survival") +
  theme(text=element_text(family="Times New Roman", size=12))
  
B1

```





