---
title: "Manuscript_GRSQ"
author: "Madeline Aberg"
date: '2022-06-13'
output: pdf_document
---

# Clean Ground Squirrel Data
## May 2, 2024


# LAST UPDATED: 9 February 2024
# USE NEGATIVE BINOMIAL OPTION!!!!!!! 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("C:/Users/Madeline/Desktop/EcologicalImpactsManuscript")

# load packages
library(nmixgof) # evaluating N-mixture models
library(tidyverse)
library(unmarked) # runs the N-mixture models
library(MuMIn) # calculates pseudo R^2
library(AICcmodavg) # gof tests (Duarte et al. 2018)
library(reshape2)
```

# Overview

## Data Collection
These data were collected by Lucian Davis and Richard Mervin in the spring/summer of 2021 following the protocol designed by Jen Cruz. At each of our paired sites, we selected a 100 m x 100 m area where we counted Piute ground squirrels. The squirrel survey areas were placed in the corner of the paired site closest to the road (but at least 50m from the road), with one exception (SF2) where the survey area was moved out of unsuitable habitat. Observers counted independent auditory and visual detections of ground squirrels along transects and for 2 minutes at each corner of the square. We waited to conduct surveys until squirrels were active above ground to maximize our chance of detecting squirrels where they were present. We conducted 3 surveys at each site between May 6, 2021 and June 14, 2021. 

## Data Analysis
Using a closed population (single-season, no migration) N-mixture analysis, which is a hierarchical model with (1) an ecological submodel that links estimated abundance to environmental predictors and (2) an observation submodel that links detection probability to predictors.

### Environmental predictors
* Recreation intensity (continuous)  
* Presence of shrub cover (binary)   

### Detectionn predictors
* Day of Year  
* Time (minutes past 12:00 AM)  
* Wind speed (km per hour)  
* Temp (degrees Celcius)  
* Observer experience level (> 1 year = experienced & < 1 year = novice)

### Analysis notes
* Combined auditory and visual detections - observers aimed to count each individual once  
* Standardized predictor variables with the equation from Gelman 2008 to be consistent with other analyses - did NOT use the unmarked package standardization  
* Used the default value for k (max observed + 100)  
* Didn't use weather as a detection predictor because there were too many categories for the number of observations and we expected that temperature and wind speed covered the most relevant aspects of the weather categories.

# Resources
This script follows the CountDataPrep.R and CountAnalysis.R scripts created by Dr. Jen Cruz, which can be accessed here: https://github.com/quantitativeconservationlab/AppPopnEco 

# Data Prep
Raw data were downloaded from the Survey123 form: https://arcg.is/CvbeX 
## Download & clean the data
```{r}
# load data 
grsqcount <- read.csv(here::here("datatemp/original/SquirrelSurvey2021.csv"))
head(grsqcount) # Each row is one survey at one site

# remove the test row -> dataset now has 30 observations (10 sites x 3 visits each)
grsqcount <- grsqcount[-1,]

# rename columns
grsqcount <- rename(grsqcount, Temp = Temperature..F.)

# create a column with total detections for each visit (combining auditory & visual)
grsqcount <- grsqcount %>%
  mutate(count = Auditory.Count + Visual.Count + Auditory.Count.1 + Visual.Count.1 + Auditory.Count.2 + Visual.Count.2 + 
           Auditory.Count.3 + Visual.Count.3 + Auditory.Count.4 + Visual.Count.4 + Auditory.Count.5 + Visual.Count.5 + 
           Auditory.Count.6  + Visual.Count.6 + Auditory.Count.7 + Visual.Count.7, na.rm = TRUE)

# Recreation intensity - values extracted from the kernel density map of all recreation points for each paired site with a 2km buffer surrounding (checked for an effect of buffer on the extraction values & there was none)
grsqcount <- grsqcount %>%
mutate(Rec = case_when(
    Site.ID=="BB1" ~ 4.426319e-08,
    Site.ID=="BB2" ~ 1.124399e-07,
    Site.ID=="BB3" ~ 9.006808e-12,
    Site.ID== "CC1" ~ 1.576085e-07, 
    Site.ID=="CC2" ~ 1.427410e-07,
    Site.ID== "PV1" ~ 1.858965e-06,
    Site.ID=="PV2" ~ 9.467036e-06,
    Site.ID== "PV3" ~ 8.465378e-07,
    Site.ID== "SF1" ~ 4.188113e-06,
    Site.ID== "SF2" ~ 3.344141e-06))

# Site number variable
grsqcount <- grsqcount %>%
  mutate(Site.num = case_when(
    Site.ID == "PV1" ~ 1, 
    Site.ID == "PV2" ~ 2, 
    Site.ID == "PV3" ~ 3, 
    Site.ID == "SF1" ~ 4, 
    Site.ID == "SF2" ~ 5, 
    Site.ID == "CC1" ~ 6, 
    Site.ID == "CC2" ~ 7, 
    Site.ID == "BB1" ~ 8, 
    Site.ID == "BB2" ~ 9, 
    Site.ID == "BB3" ~ 10))

# Date as day of year
library(lubridate)
grsqcount$date <- str_remove(grsqcount$Date, " 18:00")
grsqcount$date <- as.Date(grsqcount$date, "%m/%d/%y")
grsqcount$DOY <- yday(grsqcount$date)

# Time - want to measure as minutes past 0:00 (12AM)
grsqcount$TimeDiff <- difftime(as.POSIXct(grsqcount$Time, format = "%H:%M"), as.POSIXct("0:00", format = "%H:%M"), units = "min")

# Time - minutes past sunrise
# Sunrise from: https://www.timeanddate.com/sun/usa/boise?month=4&year=2021 
grsqcount <- mutate(grsqcount, 
                    sunrise = case_when(
                      date == "2021-05-06" ~ "06:30", 
                      date == "2021-05-13" ~ "06:21", 
                      date == "2021-05-19" ~ "06:15", 
                      date == "2021-05-20" ~ "06:14", 
                      date == "2021-05-22" ~ "06:12", 
                      date == "2021-05-24" ~ "06:11",
                      date == "2021-05-26" ~ "06:09",
                      date == "2021-05-28" ~ "06:08",
                      date == "2021-05-31" ~ "06:06",
                      date == "2021-06-03" ~ "06:05",
                      date == "2021-06-04" ~ "06:04",
                      date == "2021-06-07" ~ "06:03",
                      date == "2021-06-09" ~ "06:03",
                      date == "2021-06-14" ~ "06:02"
                    ))

grsqcount$TimeDiffSunrise <- difftime(as.POSIXct(grsqcount$Time, format = "%H:%M"), as.POSIXct(grsqcount$sunrise, format = "%H:%M"), units = "min")
grsqcount$TimeDiffSunrise <- as.numeric(grsqcount$TimeDiffSunrise)

# Temp - change units to Celcius 
library(weathermetrics)
grsqcount$Temp.Celcius <- fahrenheit.to.celsius(grsqcount$Temp)

# Wind - mph -> kmh
grsqcount$Wind.KMH <- convert_wind_speed(grsqcount$Wind, "mph", "kmph")

# Habitat - shrub as binary (0 = no shrub, 1 = shrub) * not counting Russian thistle * 
table(grsqcount$Site.ID, grsqcount$Dominant.Shrub)
grsqcount <- grsqcount %>%
  mutate(shrub.binary = case_when(
    Site.ID == "PV1" ~ 1, # Rabbitbrush
    Site.ID == "PV2" ~ 0, # Dead Russian thistle("tumbleweed")/skeleton weed
    Site.ID == "PV3" ~ 1, # Rabbitbrush  
    Site.ID == "SF1" ~ 1, # Winterfat/sagebrush
    Site.ID == "SF2" ~ 0, # Russian thistle, skeleton weed
    Site.ID == "CC1" ~ 0, # Mostly no shrub cover
    Site.ID == "CC2" ~ 0, # No shrub cover, mustard
    Site.ID == "BB1" ~ 1, # Winterfat
    Site.ID == "BB2" ~ 1, # Winterfat
    Site.ID == "BB3" ~ 0))  # No shrubs - Russian thistle

# Observer
table(grsqcount$Site.ID,grsqcount$Observer)
grsqcount <- grsqcount %>%
  mutate(obs.expr = case_when(
    grepl("Lucian", Observer) ~ 1, # 1 + years experience at the site
    grepl("RLM", Observer) ~ 2)) # less than 1 year experience at the site
grsqcount$obs.expr <- grsqcount$obs.expr %>% replace(is.na(.), 1) # adding my surveys to the experienced category

# Add % shrub from bird pt counts 
grsqcount <- grsqcount %>%
  mutate(
    shrub.per = case_when(    # all pts
      Site.ID == "BB1" ~ 1.25,
      Site.ID == "BB2" ~ 3.50,
      Site.ID == "BB3" ~ 0.00, 
      Site.ID == "CC1" ~ 2.38,
      Site.ID == "CC2" ~ 5.38,
      Site.ID == "PV1" ~ 7.13,
      Site.ID == "PV2" ~ 6.38,
      Site.ID == "PV3" ~ 11.56,
      Site.ID == "SF1" ~ 2.44,
      Site.ID == "SF2" ~ 7.63), 
    shrub.pt = case_when(
      Site.ID == "BB1" ~ 5,
      Site.ID == "BB2" ~ 10,
      Site.ID == "BB3" ~ 0, 
      Site.ID == "CC1" ~ 1,
      Site.ID == "CC2" ~ 0,
      Site.ID == "PV1" ~ 10,
      Site.ID == "PV2" ~ 1,
      Site.ID == "PV3" ~ 5,
      Site.ID == "SF1" ~ 0,
      Site.ID == "SF2" ~ 30))

# Check predictor structure
str(grsqcount$Site.ID)
str(grsqcount$Survey.Day)
str(grsqcount$obs.expr)
grsqcount$obs.expr <- as.factor(grsqcount$obs.expr)
str(grsqcount$Wind.KMH) 
str(grsqcount$Temp.Celcius) 
str(grsqcount$DOY)
str(grsqcount$TimeDiff)
grsqcount$TimeDiff <- as.numeric(grsqcount$TimeDiff)
str(grsqcount$Rec)
str(grsqcount$shrub.binary)
grsqcount$shrub.binary <- as.factor(grsqcount$shrub.binary)
str(grsqcount$count)
str(grsqcount$shrub.per)
str(grsqcount$shrub.pt)
```

# Save clean dataset
```{r}
write.csv(grsqcount, "C:/Users/Madeline/Desktop/EcologicalImpactsManuscript/CleanData/GroundSquirrel_Clean_29April2024")
```

