---
title: "Manuscript_RaptorRaven"
author: "Madeline Aberg"
date: '2022-06-13'
output: pdf_document
---
# Clean Raptor & Raven Data
# 29 Apr 2024

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(ggplot2)
library(jtools)
library(lubridate)
```

# Overview

## Data Collection
These data were collected by Lucian Davis in the spring/summer of 2021 following the protocol designed by Lucian Davis, Rob Miller and Jay Carlisle. At each of our paired sites, we counted raptors, ravens, and mammalian scavengers on 8 survey days. Observers followed the same transect through 16 evenly spaced (4x4 grid, 250m between points) during each survey. Observers counted independent auditory and visual detections of scavengers along along transects. We conducted 8 surveys at each site between May 12, 2021, and July 9, 2021, aiming to survey each site at various times of day. 

## Data Analysis
We used a closed population (single-season, no migration) N-mixture analysis, which is a hierarchical model with (1) an ecological submodel that links abundance to environmental predictors and (2) an observation submodel that links detection probability to predictors. We created one model for raven abundance and one model for raptor abundance. The raptor abundance model includes species as a random effect. 
  
### Environmental predictors
* Recreation intensity (continuous)   
* % shrub cover (continuous)  
* Ground squirrel abundance estimate (continuous) *removed because of high multicolinearity with power line density*   
* Road density (continuous)  *removed because of high multicolinearity with % developed*  
* Power line density (continuous)  
* % developed (continuous)  

### Detection predictors
* Day of Year  
* Time (minutes past 12:00 AM)  
* Wind speed (km per hour)  
* Temp (Celsius)  

### Analysis notes
* Combined auditory and visual detections - observers aimed to count each individual once   
* Removed observations out of the buffer  
* Combined all Buteo species, including unknown Buteo  
* Standardized predictor variables with the equation from Gelman 2008 to be consistent with other analyses - did NOT use the unmarked package standardization  
* Used the default value for k (max observed + 100)  
* An assumption of N-Mixture models is that all observations are individuals, no groups. Our data breaks this assumption - the dataset with observations in the buffer has 94 observations with >1 (74 x 2 birds, 10 x 3 birds, 8 x 4 birds, 2 x 5 birds). To be conservative, I'm counting these as only 1 individual, since we don't have a way to account for higher detection probability for more individuals (because observations are aggregated by survey day, so we lose some individual info).   
* Didn't use weather as a detection predictor because there were too many categories for the number of observations and we expected that temperature and wind speed covered the most relevant aspects of the weather categories. 

# Resources
This analysis follows the CountDataPrep.R and CountAnalysis.R scripts created by Dr. Jen Cruz, which can be accessed here: https://github.com/quantitativeconservationlab/AppPopnEco. 
## UBMS package
https://cran.r-project.org/web/packages/ubms/vignettes/random-effects.html.
https://groups.google.com/g/unmarked/c/dChRYcvox6U?pli=1 
https://stats.stackexchange.com/questions/31569/questions-about-how-random-effects-are-specified-in-lmer *specifying random effects*
https://kenkellner.com/blog/ubms-vignette.html 


# Data Prep
## Load & combine datasets
```{r}
# Load the data (downloaded from Survey123)
RS1 <- read.csv(here::here("datatemp/original/RaptorSurveys2021/RaptorSurvey_1.0.csv"))
RS1_scav <- read.csv(here::here("datatemp/original/RaptorSurveys2021/RaptorSurvey_1.0_Scavenger.csv"))
RS2 <- read.csv(here::here("datatemp/original/RaptorSurveys2021/RaptorSurvey_2.0.csv"))
RS2_scav <- read.csv(here::here("datatemp/original/RaptorSurveys2021/RaptorSurvey_2.0_Scavenger.csv"))
RS3 <- read.csv(here::here("datatemp/original/RaptorSurveys2021/RaptorSurvey_3.0.csv"))
RS3_scav <- read.csv(here::here("datatemp/original/RaptorSurveys2021/RaptorSurvey_3.0_Scavenger.csv"))

# select the variables that we will use 
RS1 <- RS1 %>%
  dplyr::select(GlobalID, Date, Start.Time, Observer,Site.Name, Temperature..F., Wind.Speed, Weather, Number.of.Recreation.Groups, End.Time, Notes)

RS2 <- RS2 %>%
  dplyr::select(GlobalID, Date, Start.Time, Observer, Site.Name, Temperature..F., Wind.Speed, Weather, Number.of.Recreation.Groups, End.Time, Notes)

RS3 <- RS3 %>%
  dplyr::select(GlobalID, Date, Start.Time, Observer, Site.Name, Temperature..F., Wind.Speed, Weather, Number.of.Recreation.Groups, End.Time, Notes)

RS1_scav <- RS1_scav %>%
  dplyr::select(Species, Count, Location, ParentGlobalID)

RS2_scav <- RS2_scav %>%
  dplyr::select(Species, Count, Location, ParentGlobalID)

RS3_scav <- RS3_scav %>%
  dplyr::select(Species, Count, Location, ParentGlobalID)

## combine the scavenger observation datasets
RaptorObs <- full_join(RS1_scav, RS2_scav) 
RaptorObs <- full_join(RaptorObs, RS3_scav)

## combine survey datasets
SurveyInfo <- full_join(RS1, RS2) 
SurveyInfo <- full_join(SurveyInfo, RS3) 
```

## Add site-level variables to SurveyInfo
```{r}
# Ground Squirrel estimates from Manuscript_GRSQ.Rmd
SurveyInfo <- SurveyInfo %>%
  mutate(grsq_est = case_when(
    Site.Name == "BB1" ~ 81.3,
    Site.Name == "BB2" ~ 58.1,
    Site.Name == "BB3" ~ 15.9, 
    Site.Name == "CC1" ~ 70.6,
    Site.Name == "CC2" ~ 112.5,
    Site.Name == "PV1" ~ 15.5,
    Site.Name == "PV2" ~ 63.7,
    Site.Name == "PV3" ~ 63.1,
    Site.Name == "SF1" ~ 15.8,
    Site.Name == "SF2" ~ 24.7))

# Adding road density -  from a shapefile from the Idaho Army National Guard & calculated in QGIS 
SurveyInfo <- SurveyInfo %>%
  mutate(roads = case_when(
    Site.Name == "BB1" ~ 0,
    Site.Name == "BB2" ~ 14,
    Site.Name == "BB3" ~ 7, 
    Site.Name == "CC1" ~ 29,
    Site.Name == "CC2" ~ 15,
    Site.Name == "PV1" ~ 28,
    Site.Name == "PV2" ~ 1,
    Site.Name == "PV3" ~ 0,
    Site.Name == "SF1" ~ 0,
    Site.Name == "SF2" ~ 0))

# % shrub cover - from point counts in April 2021 - includes all shrubs
SurveyInfo <- SurveyInfo %>%
  mutate(shrub = case_when(
    Site.Name == "BB1" ~ 1.25,
    Site.Name == "BB2" ~ 3.50,
    Site.Name == "BB3" ~ 0.00, 
    Site.Name == "CC1" ~ 2.38,
    Site.Name == "CC2" ~ 5.38,
    Site.Name == "PV1" ~ 7.13,
    Site.Name == "PV2" ~ 6.38,
    Site.Name == "PV3" ~ 11.56,
    Site.Name == "SF1" ~ 2.44,
    Site.Name == "SF2" ~ 7.63))

# % developed - from NLCD 2016 analyzed in QGIS - using all levels of developed
SurveyInfo <- SurveyInfo %>%
  mutate(developed = case_when(
    Site.Name == "BB1" ~ 0,
    Site.Name == "BB2" ~ 0,
    Site.Name == "BB3" ~ 0, 
    Site.Name == "CC1" ~ 2.75,
    Site.Name == "CC2" ~ 0,
    Site.Name == "PV1" ~ 3.57,
    Site.Name == "PV2" ~ 0,
    Site.Name == "PV3" ~ 0,
    Site.Name == "SF1" ~ 0,
    Site.Name == "SF2" ~ 0))

# Recreation - recreation intensity (using kernel density of all rec points within a site from March - July 2021)
SurveyInfo <- SurveyInfo %>%
  mutate(rec.intensity = case_when(
    Site.Name=="BB1" ~ 4.426319e-08,
    Site.Name=="BB2" ~ 1.124399e-07,
    Site.Name=="BB3" ~ 9.006808e-12,
    Site.Name== "CC1" ~ 1.576085e-07, 
    Site.Name=="CC2" ~ 1.427410e-07,
    Site.Name== "PV1" ~ 1.858965e-06,
    Site.Name=="PV2" ~ 9.467036e-06,
    Site.Name== "PV3" ~ 8.465378e-07,
    Site.Name== "SF1" ~ 4.188113e-06,
    Site.Name== "SF2" ~ 3.344141e-06))

# Powerline density - calculated in QGIS from shapefiles of powerline locations that we collected in the field
SurveyInfo <- SurveyInfo %>%
  mutate(powerline = case_when(
    Site.Name == "BB1" ~ 2.33,
    Site.Name == "BB2" ~ 3.35,
    Site.Name == "BB3" ~ 2.2, 
    Site.Name == "CC1" ~ 0,
    Site.Name == "CC2" ~ 0,
    Site.Name == "PV1" ~ 4.96,
    Site.Name == "PV2" ~ 0,
    Site.Name == "PV3" ~ 0,
    Site.Name == "SF1" ~ 5.45,
    Site.Name == "SF2" ~ 5.74))
```

## Clean RaptorObs and SurveyInfo -> RaptorFull (1 row per scavenger observation with obs + survey data)
```{r, include=FALSE}

#### RaptorObs ####
RaptorObs <- rename(RaptorObs, GlobalID = ParentGlobalID)  # rename so that it matches the SurveyInfo dataset

# filter out observations out of the buffer 
RaptorInBuffer <- RaptorObs %>%
  filter(., Location != "out") 

# Replace NA values in the count to be 1 
RaptorInBuffer$Count <- RaptorInBuffer$Count %>% replace(is.na(.), 1)

#### SurveyInfo ####

# Fix Observer name
SurveyInfo$Observer <- dplyr::recode(SurveyInfo$Observer, "Lucian " = "Lucian")

#### Rename variables ####
SurveyInfo <- rename(SurveyInfo, Temp = Temperature..F.)

# Site-level predictors
str(SurveyInfo$Site.Name) # character, want to be as factor
SurveyInfo$Site.Name <- as.factor(SurveyInfo$Site.Name)
str(SurveyInfo$grsq) # predictor variable - numerical = good!  
str(SurveyInfo$rec.intensity) # numerical 
str(SurveyInfo$shrub) # predictor - num
str(SurveyInfo$roads) # predictor - num
str(SurveyInfo$developed) # predictor - num
str(SurveyInfo$powerline) # predictor - num

# Observation-level predictors
str(SurveyInfo$Weather) # chr, want to be a factor
SurveyInfo$Weather <- as.factor(SurveyInfo$Weather)
str(SurveyInfo$Date) # want to read as date & remove 18:00 from the label (not correct time)
SurveyInfo$Date <- str_remove(SurveyInfo$Date, " 18:00")
SurveyInfo$Date <- as.Date(SurveyInfo$Date, "%m/%d/%y") # now want to be day of year
SurveyInfo$DayOfYear <- yday(SurveyInfo$Date) # Date as day of year
# Time - want to change to minutes past 0:00
SurveyInfo$TimeDiff <- difftime(as.POSIXct(SurveyInfo$Start.Time, format = "%H:%M"), as.POSIXct("0:00", format = "%H:%M"), units = "min")
SurveyInfo$TimeDiff <- as.numeric(SurveyInfo$TimeDiff)

# Time - minutes past sunrise https://www.timeanddate.com/sun/usa/boise?month=5&year=2021
table(SurveyInfo$Date)
SurveyInfo <- mutate(SurveyInfo, 
                    sunrise = case_when(
                      Date == "2021-05-05" ~ "06:31",
                      Date == "2021-05-06" ~ "06:30", 
                      Date == "2021-05-12" ~ "06:22",
                      Date == "2021-05-13" ~ "06:21", 
                      Date == "2021-05-19" ~ "06:15", 
                      Date == "2021-05-20" ~ "06:14", 
                      Date == "2021-05-22" ~ "06:12", 
                      Date == "2021-05-24" ~ "06:11",
                      Date == "2021-05-25" ~ "06:10",
                      Date == "2021-05-26" ~ "06:09",
                      Date == "2021-05-28" ~ "06:08",
                      Date == "2021-05-31" ~ "06:06",
                      Date == "2021-06-01" ~ "06:06",
                      Date == "2021-06-04" ~ "06:04",
                      Date == "2021-06-05" ~ "06:04",
                      Date == "2021-06-07" ~ "06:03",
                      Date == "2021-06-08" ~ "06:03",
                      Date == "2021-06-09" ~ "06:03",
                      Date == "2021-06-10" ~ "06:03",
                      Date == "2021-06-11" ~ "06:03",
                      Date == "2021-06-12" ~ "06:02",
                      Date == "2021-06-14" ~ "06:02",
                      Date == "2021-06-15" ~ "06:02",
                      Date == "2021-06-16" ~ "06:02",
                      Date == "2021-06-17" ~ "06:02",
                      Date == "2021-06-18" ~ "06:03",
                      Date == "2021-06-20" ~ "06:03",
                      Date == "2021-06-21" ~ "06:03",
                      Date == "2021-06-22" ~ "06:03",
                      Date == "2021-06-23" ~ "06:04",
                      Date == "2021-06-28" ~ "06:06",
                      Date == "2021-07-01" ~ "06:07",
                      Date == "2021-07-02" ~ "06:08",
                      Date == "2021-07-03" ~ "06:08",
                      Date == "2021-07-04" ~ "06:09",
                      Date == "2021-07-05" ~ "06:09",
                      Date == "2021-07-06" ~ "06:10",
                      Date == "2021-07-07" ~ "06:11",
                      Date == "2021-07-08" ~ "06:12",
                      Date == "2021-07-09" ~ "06:12",
                      Date == "2021-07-12" ~ "06:15",
                      Date == "2021-07-13" ~ "06:15"))

SurveyInfo$TimeDiffSunrise <- difftime(as.POSIXct(SurveyInfo$Start.Time, format = "%H:%M"), as.POSIXct(SurveyInfo$sunrise, format = "%H:%M"), units = "min")
SurveyInfo$TimeDiffSunrise <- as.numeric(SurveyInfo$TimeDiffSunrise)

library(weathermetrics)
str(SurveyInfo$Wind.Speed)
SurveyInfo$Wind.Speed <- SurveyInfo$Wind.Speed%>%replace(is.na(.), 5) 
SurveyInfo$Wind.KMHR <- convert_wind_speed(SurveyInfo$Wind.Speed, "mph", "kmph")
str(SurveyInfo$Temp)
SurveyInfo$Temp <- as.numeric(SurveyInfo$Temp)
SurveyInfo$Temp.Celcius <- fahrenheit.to.celsius(SurveyInfo$Temp)

SurveyInfo <- dplyr::select(SurveyInfo, c(GlobalID, Observer, Site.Name, grsq_est, roads, shrub, developed, rec.intensity, powerline, DayOfYear, TimeDiff, TimeDiffSunrise, Wind.KMHR, Temp.Celcius))
```






## Dataframe for analysis: RavenSurveys - 1 row per survey day with summary of # ravens *** Using observations in the buffer only ***
```{r}
# Now want to have a dataframe with each survey day as the row and summaries of the total count, species counts as columns - could do this before the first join ***
RavenSurveys <- RaptorInBuffer %>%
  group_by(GlobalID)%>%
  summarize(RavenCount = sum(Count[Species=="CORA"],na.rm = TRUE))

RavenSurveys <- full_join(RavenSurveys, SurveyInfo, by = "GlobalID", keep = TRUE) # This is our dataframe with each survey day as a row - this is what we'll use for the analysis 

tibble(RavenSurveys)
```

# Clean RavenSurveys dataframe
```{r}
# Remove test surveys
RavenSurveys <-dplyr::filter(RavenSurveys, Observer != "Update TEST " & Observer != "TEST" & Observer != "Test") # should remove 3 obs


### Fix NAs ####
# There are two surveys (BB3 & SF2) that had 0 observations of birds within the buffer, setting those counts = 0 
RavenSurveys$RavenCount <- RavenSurveys$RavenCount %>% replace(is.na(.), 0)

str(RavenSurveys$RavenCount)

```

## Save Clean Raven Surveys
```{r}
write.csv(RavenSurveys, "C:/Users/Madeline/Desktop/EcologicalImpactsManuscript/CleanData/Ravens_Clean_29April2024.csv")
```




# Clean RaptorSurveys dataframe
```{r}
# Now want to have a dataframe with each survey day as the row and summaries of the total count, species counts as columns - could do this before the first join ***

RaptorSurveys <- RaptorInBuffer %>%
  group_by(GlobalID)%>%
  summarize(
    a =  sum(Count[Species=="BUOW"],na.rm = TRUE), 
    b = sum(Count[Species=="buteo"],na.rm = TRUE), 
    c = sum(Count[Species=="FEHA"],na.rm = TRUE), 
    d = sum(Count[Species=="GOEA"],na.rm = TRUE),
    e = sum(Count[Species=="NOHA"],na.rm = TRUE), 
    f = sum(Count[Species=="PRFA"],na.rm = TRUE), 
    g = sum(Count[Species=="RTHA"],na.rm = TRUE), 
    h = sum(Count[Species=="SWHA"],na.rm = TRUE), 
    RaptorCount = a + b + c + d + e + f + g + h)

RaptorSurveys <- full_join(RaptorSurveys, SurveyInfo, by = "GlobalID", keep = TRUE) # This is our dataframe with each survey day as a row - this is what we'll use for the analysis 

tibble(RaptorSurveys)
```

```{r}
# Remove test surveys
RaptorSurveys <-dplyr::filter(RaptorSurveys, Observer != "Update TEST " & Observer != "TEST" & Observer != "Test") # should remove 3 obs


### Fix NAs ####
# There are two surveys (BB3 & SF2) that had 0 observations of birds within the buffer, setting those counts = 0 
RaptorSurveys$RaptorCount <- RaptorSurveys$RaptorCount %>% replace(is.na(.), 0)

str(RaptorSurveys$RaptorCount)
```
## Save Clean Raptor Surveys
```{r}
write.csv(RaptorSurveys, "C:/Users/Madeline/Desktop/EcologicalImpactsManuscript/CleanData/Raptors_Clean_29April2024.csv")
```

# Save RaptorInBuffer
```{r}
write.csv(RaptorInBuffer, "C:/Users/Madeline/Desktop/EcologicalImpactsManuscript/CleanData/RaptorsInBuffer_Clean_29April2024.csv")
```

